---
title: "Report"
author: "Yang Xia"
date: "2017/11/13"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Specifications of host execution Environment
Attribute                         | Value
----------------------------------|-------------------------------
Java Version                      | 1.8.0_102
Spark Version                     | 2.2.0
Model Identifier                  |	MacBookPro13,2
Processor Name                    |	Intel Core i5
Processor Speed                   |	2.2 GHz
Number of Processors              |	1
Total Number of Cores             |	4
Memory                            | 16 GB
Driver Memory                     | 2 GB
Executor Memory                   | 2 GB

### Summary of the design

The implementation start with reading the data. I had a class called "SongInfo" which store the related information as fileds. This makes the code cleaner. I didn't do the so-called clean-up test before hand due to the fact that this might cause inaccuracy in the result. The milionSong object is the entry point to the program that taken in song_info.csv and artist_terms.csv locations in the input file and calls the methods in the SongInfo class to get the desired output. 

## Result Analysis

####Assumptions
Even if one of the cell is empty or doesn't have the correct value, the other part of the record is invalid.

For the whole dataset:

#### Number of distinct songs: 999056

#### Number of distinct artists: 44745

#### Number of distinct albums: 149275

#### Top 5 loudest songs: 
(Assumption: a larger loudness implies a louder song)

Song Id                         | Song Name                           | Loudness Score
--------------------------------|-------------------------------------|------------------------------------
SOCMYZF12AB0186FF4              | Modifications                       | 4.318
SONHDXQ12AB018C1F1              | Track 02                            | 4.3
SOZCIHV12AC46894E3              | Hey You Fuxxx!                      | 4.231
SOEJMJF12AC90715D1              | War Memorial Exit                   | 4.166
SOBEMPS12A8C13BD46              | Meta Abuse                          | 4.15

#### Top 5 longest songs:
(Assumption: a larger duration implies a longer song)

Song Id                         | Song Name                               | Duration
--------------------------------|-----------------------------------------|--------------------
SOOUBST12AC90977B6              | Grounation                              | 3034.90567
SOXUCQN12A6D4FC451              | Raag - Shuddha KalyaN                   | 3033.5995
SOOMVZJ12AB01878EB              | Discussion 2                            | 3033.44281
SOTNVEE12A8C13F470              | Chapitre Un (a): Toutes Les Histoires   | 3032.76363
SOGFXNB12A8C137BE5              | Der Geist des Llano Estacado  Ein Spion | 3032.58077

#### Top 5 fastest songs:
(Assumption: a larger  tempo implies a faster song)

Song Id                         | Song Name                       | Tempo Score
--------------------------------|---------------------------------|------------------------------------
SOVVTEZ12AB0184AAB              | Beep Beep                       | 302.3
SOMSJWX12AB017DB99              | Late Nite Lounge: WVIP          | 296.469
SOUTBKH12A8C136286              | A Place Called Hope             |285.157
SOEVQJB12AC960DA2C              | Bellas Lullaby - Perrier Citron |284.208
SOTUXOB12AB0188C3A              | Troubled Times                  |282.573


#### Top 5 most familiar artists:
Artist Name                        | Artist Id                      | Familiarity Score
-----------------------------------|--------------------------------|------------------------------------
Akon                                | ARCGJ6U1187FB4D01F            | 1.0
Akon_ San Quinn_ JT the Bigga Figga | ARCGJ6U1187FB4D01F            | 1.0
Akon / Eminem                       | ARCGJ6U1187FB4D01F            | 1.0
Akon / Styles P                     | ARCGJ6U1187FB4D01F            | 1.0
Akon / Wyclef Jean                  | ARCGJ6U1187FB4D01F            | 1.0

#### Top 5 hottest artists:
Artist Name                        | Artist Id                      | Hotness Score
-----------------------------------|--------------------------------|------------------------------------
Daft Punk                          | ARF8HTQ1187B9AE693             | 0.997066533839045
Daft Punk                          | ARF8HTQ1187B9AE693             | 0.997004803235357
Black Eyed Peas                    | ARTDQRC1187FB4EFD4             | 0.982623202516712
Kanye West                         | ARRH63Y1187FB47783             | 0.972399563931911
Kanye West / Jamie Foxx            | ARRH63Y1187FB47783             | 0.972399563931911


##### Top 5 hottest songs:
Song Id                        | Song Name                          | Hotness Score
-------------------------------|------------------------------------|------------------------
SONMVZB12AB01829BA             | Der Maggot Tango                   | 0.521321041187445
SOAGFNE12A8C134D82             | Donde Caigo                        | 0.454192988218022
SOIUPEW12A8C13BCB2             | Willow Weep For Me (Live)          | 0.266955186275539
SOAYGMO12A6D4F69E8             | I'd Have Never Found Somebody New  | 0.249065794853703
SOFJZNE12A8AE45C1F             | The??blind Walk Over The Edge       | 0.492398352817721


#### Top 5 hottest genres (mean artists hotness in artist_term)
Genres                                       | Mean Hotness                    
---------------------------------------------|------------------------------
christmas song                               | 0.6084021       
kotekote                                     | 0.6022205
female artist                                | 0.5753481
girl rockers                                 | 0.5737886    
alternative latin                            | 0.5737588
 


##### Top 5 Prolific artists:
Artist Id                        | Artist Name                  | Number of songs by the artist
-----------------------------------|----------------------------|------------------------------------
AR6681Y1187FB39B02                  | Ike & Tina Turner         | 208
ARXPPEY1187FB51DF4                  | Michael Jackson           |204
ARH861H1187B9B799E                  | Johnny Cash               | 201
AR8L6W21187B9AD317                  | Diana Ross & The Supremes | 196
ARLHO5Z1187FB4C861                  | Beastie Boys              | 194

#### Top 5 most popular keys (must have confidence > 0.7):

Key                        | Ocurrances                    
-----------------------------------|------------------------------
7                         | 53144       
9                         | 46855          
2                         | 46431     
4                         | 35099
11                        | 35078       

#### Top 5 most common words in song titles (excluding articles, prepositions, conjunctions):
(Assumption: (ablum version) doesn't count as part of the title)

Word                        | Ocurrances                    
-----------------------------------|------------------------------
LOVE                        | 9882       
LIVE                        | 6192          
YOUR                        | 5333     
NO                          | 4822
REMIX                       | 4462 

### Performance Analysis

I run the programm on the whole data set a couple of times, and in average it takes around 3 minutes to run on the whole dataset. Instead of ploting the volin graph for the total runing time, I think it's more intersting to look at the time spent on every query, and compare the performance with and withour persist.

####Times taken for different operations on **entire data with persist**:
```{r echo=FALSE}
library(ggplot2)
tasksMap = read.csv("output/TasksTable", header = FALSE)
colnames(tasksMap) <- c('Task','Query performed')
print(tasksMap)
timesAll = read.csv("output/Time Counter on whole data",header = FALSE)
colnames(timesAll) <- c('Task','Time')
total_time <- timesAll[nrow(timesAll),2]
timesAll <- timesAll[1:nrow(timesAll)-1,]
timesAll <- timesAll[order(timesAll$Task),]
plot5 <- ggplot(timesAll, aes(x=Task,
                           y=timesAll$Time,
                           fill=Task)) + geom_bar(stat = 'identity') + labs(title="Execution Times for each Task on Entire Data", x="Task", y="Time in millisecs", fill="Task")
plot5

```
####Times taken for different operations on the **entire data without persist**:

```{r echo=FALSE}
library(ggplot2)
timesWithoutPersist = read.csv("output/Time Counter without persist",header = FALSE)
colnames(timesWithoutPersist) <- c('Task','Time')
total_time <- timesWithoutPersist[nrow(timesWithoutPersist),2]
timesWithoutPersist <- timesWithoutPersist[1:nrow(timesWithoutPersist)-1,]
timesWithoutPersist <- timesWithoutPersist[order(timesWithoutPersist$Task),]
plot4 <- ggplot(timesWithoutPersist, aes(x=Task,
                           y=timesWithoutPersist$Time,
                           fill=Task)) + geom_bar(stat = 'identity')+ labs(title="Execution Times for each Task on Entire Data without Persisting", x="Task", y="Time in millisecs", fill="Task")
plot4
```

I ran the code on the subset as well as the entire dataset. Two things that are interesting.
First, We can see that for majority of the tasks, the execution time is faster when we do not persist the RDD and keep generating it on the fly. This is because we are running our code in a standalone mode rather than in a distributed mode.
Second, the "hottest genres" query takes up most of the time. I think it's because a joining two is very expensive. Since it's a bottleneck in this program, we can try to optimize it in future work.

### Conclusion

This program takes million songs data as input and do some simple analysis on the whole dataset. Using Spark instead of Hadoop saves a lot of effort. Though I've never implemented in hadoop, I can imagine doing the same things in Hadoop require many map job and reduce job, which in most of the cases are not as efficient as I did here with spark. 
The bottleneck here is the join operation; some optimization could be done to improve the performance.
In a stand-alone mod, persist does not give me the speed-up as it should be while in a distributed mode.
