---
title: "A7 - Report"
author: "Ritvika Reddy, Yang Xia"
date: "11/3/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Local Machine Specifications:
```
  Processor Name:	Intel Core i5
  Processor Speed:	1.6 GHz
  Number of Processors:	1
  Total Number of Cores:	2
  L2 Cache (per Core):	256 KB
  L3 Cache:	3 MB
  Memory:	8 GB
```

#### Implementing K-Means on `Subset Data` for `k=3` in `10 iterations` for:

```{r echo=FALSE}
loudness_clusters <- read.csv("output/subset/loudness.csv", quote = "")
length_clusters <- read.csv("output/subset/length.csv", quote = "")
tempo_clusters <- read.csv("output/subset/tempo.csv", quote = "")
hotness_clusters <- read.csv("output/subset/hotness.csv", quote = "")
combinedHotness_clusters <- read.csv("output/subset/artist_hotness_hotness.csv",quote = "")

```

1. Loudness
---- Cluster with centroid -15.27 is Quiet , -9.59 is Medium and -5.93 is Loud.
```{r echo=FALSE}
library(ggplot2)

plot1 <- ggplot(loudness_clusters, aes(x=loudness_clusters$loudness_x,
                                        y=c(0))
                 ) + geom_point(aes(color=(loudness_clusters$Centroid_x))) + labs(x="Loudness", y="", title = "Clustering Loudness", color="Centroids") + theme_classic()
plot1

plot2 <- ggplot(loudness_clusters, aes(x=loudness_x,fill=factor(loudness_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 1)  + labs(x="Loudness", y="", title = "Clustering Loudness", fill="Centroids") + theme_classic()
plot2

```
2. Duration
```{r echo=FALSE}
plot3 <- ggplot(length_clusters, aes(x=length_clusters$length_x,
                                        y=c(0))
                 ) + geom_point(aes(color=(length_clusters$Centroid_x))) + labs(x="Duration", y="", title = "Clustering Duration", color="Centroids") + theme_classic()
plot3

plot4 <- ggplot(length_clusters, aes(x=length_x,fill=factor(length_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 10)  + labs(x="Duration", y="", title = "Clustering Duration", fill="Centroids") + theme_classic()
plot4

```
3. Tempo
```{r echo=FALSE}
plot5 <- ggplot(tempo_clusters, aes(x=tempo_clusters$tempo_x,
                                        y=c(0))
                 ) + geom_point(aes(color=(tempo_clusters$Centroid_x))) + labs(x="Tempo", y="", title = "Clustering Tempo", color="Centroids") + theme_classic()
plot5

plot6 <- ggplot(tempo_clusters, aes(x=tempo_x,fill=factor(tempo_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 3.5)  + labs(x="Tempo", y="", title = "Clustering Tempo", fill="Centroids") + theme_classic()
plot6

```
4. Song Hotness
```{r echo=FALSE}
plot7 <- ggplot(hotness_clusters, aes(x=hotness_clusters$hotness_x,
                                        y=c(0))
                 ) + geom_point(aes(color=(hotness_clusters$Centroid_x))) + labs(x="Song Hotness", y="", title = "Clustering Hotness", color="Centroids") + theme_classic()
plot7

plot8 <- ggplot(hotness_clusters, aes(x=hotness_x,fill=factor(hotness_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 0.025)  + labs(x="Hotness", y="", title = "Clustering Hotness", fill="Centroids") + theme_classic()
plot8

```

5. Combined Hotness
```{r echo=FALSE}
plot9 <- ggplot(combinedHotness_clusters, aes(x=combinedHotness_clusters$artist_hotness_x,
                                        y=combinedHotness_clusters$hotness_y)
                 ) + geom_point(aes(color=(combinedHotness_clusters$Centroid_x))) + labs(x="Artist Hotness", y="Song Hotness", title = "Clustering Combined Hotness", color="Centroids") + theme_classic()
plot9

plot10 <- ggplot(combinedHotness_clusters, aes(x=combinedHotness_clusters$artist_hotness_x, y=combinedHotness_clusters$hotness_y,fill=factor(hotness_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 0.05,stackdir="center")  + labs(x="Artist Hotness", y="Song Hotness", title = "Clustering Combined Hotness", fill="Centroids") + theme_classic()
plot10

```
#### Implementing K-Means on `Entire Data` for `k=3` in `10 iterations` for:

```{r echo=FALSE}
all_loudness_clusters <- read.csv("output/all//loudness.csv", quote = "",strip.white = TRUE,na.string = "")
all_length_clusters <- read.csv("output/all/length.csv", strip.white = TRUE,quote = "",na.string = "")
all_tempo_clusters <- read.csv("output/all/tempo.csv", strip.white = TRUE,quote = "",na.string = "")
all_hotness_clusters <- read.csv("output/all/hotness.csv", strip.white = TRUE,quote = "",na.string = "")
all_combinedHotness_clusters <- read.csv("output/all/artist_hotness_hotness.csv",strip.white = TRUE,quote = "",na.string = "")

```

1. Loudness
```{r echo=FALSE, fig.width=20}
library(ggplot2)
all_loudness_clusters <- all_loudness_clusters[complete.cases(all_loudness_clusters),] 
# all_loudness_clusters[-c(1222588,161586,104675,352835,323343,31604)]
plot11 <- ggplot(all_loudness_clusters, aes(x=all_loudness_clusters$loudness_x,
                                        y=c(0))
                 ) + geom_point(aes(color=(all_loudness_clusters$Centroid_x))) + labs(x="Loudness", y="", title = "Clustering Loudness", color="Centroids") + theme_classic()
plot11
```
```{r echo=FALSE}
# plot12 <- ggplot(all_loudness_clusters, aes(x=loudness_x,fill=factor(all_loudness_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 1)  + labs(x="Loudness", y="", title = "Clustering Loudness", fill="Centroids") + theme_classic()
# plot12
```
2. Duration
```{r echo=FALSE, fig.width=15}
plot13 <- ggplot(all_length_clusters[-c(171877,358856,360492,335677,312704)], aes(x=all_length_clusters$length_x,
                                        y=c(0))
                 ) + geom_point(aes(color=(all_length_clusters$Centroid_x))) + labs(x="Duration", y="", title = "Clustering Duration", color="Centroids") + theme_classic()
plot13

# plot14 <- ggplot(all_length_clusters, aes(x=all_length_x,fill=factor(all_length_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 10)  + labs(x="Duration", y="", title = "Clustering Duration", fill="Centroids") + theme_classic()
# plot14
```
3. Tempo
```{r echo=FALSE, fig.width=15}
plot15 <- ggplot(all_tempo_clusters, aes(x=all_tempo_clusters$tempo_x,
                                        y=c(0))
                 ) + geom_point(aes(color=(all_tempo_clusters$Centroid_x))) + labs(x="Tempo", y="", title = "Clustering Tempo", color="Centroids") + theme_classic()
plot15

#plot16 <- ggplot(all_tempo_clusters, aes(x=all_tempo_x,fill=factor(all_tempo_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 3.5)  + labs(x="Tempo", y="", title = "Clustering Tempo", fill="Centroids") + theme_classic()
#plot16
```
4. Song Hotness
```{r echo=FALSE, fig.width=15}
plot17 <- ggplot(all_hotness_clusters, aes(x=all_hotness_clusters$hotness_x,
                                        y=c(0))
                 ) + geom_point(aes(color=(all_hotness_clusters$Centroid_x))) + labs(x="Song Hotness", y="", title = "Clustering Hotness", color="Centroids") + theme_classic()
plot17

#plot18 <- ggplot(all_hotness_clusters, aes(x=all_hotness_x,fill=factor(all_hotness_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 0.025)  + labs(x="Hotness", y="", title = "Clustering Hotness", fill="Centroids") + theme_classic()
#plot18
```

5. Combined Hotness
```{r echo=FALSE, fig.width=15}
plot19 <- ggplot(all_combinedHotness_clusters, aes(x=all_combinedHotness_clusters$artist_hotness_x,
                                        y=all_combinedHotness_clusters$hotness_y)
                 ) + geom_point(aes(color=(all_combinedHotness_clusters$Centroid_x))) + labs(x="Artist Hotness", y="Song Hotness", title = "Clustering Combined Hotness", color="Centroids") + theme_classic()
plot19

#plot20 <- ggplot(all_combinedHotness_clusters, aes(x=all_combinedHotness_clusters$artist_hotness_x, y=all_combinedHotness_clusters$hotness_y,fill=factor(all_hotness_clusters$Centroid_x))) + geom_dotplot(dotsize = 0.5, binwidth = 0.05)  + labs(x="Artist Hotness", y="Song Hotness", title = "Clustering Combined Hotness", fill="Centroids") + theme_classic()
#plot20
```

#### Performance Times on Subset and Entire Data
```{r echo=FALSE}
cat("Time Taken for K-Means on each Metric for Entire Data in millisecs" , "\n")
cat(readLines("output/all/Time_counter.txt"), sep = "\n")
cat("\n\n")
cat("Time Taken for K-Means on each Metric for Subset Data in millisecs" , "\n")
cat(readLines("output/subset//Time_counter"), sep = "\n")

```


#### Conclusions:
1. We are performing fuzzy clustering because in practical sense, k-means algorithm doesn't converge in 10 iterations for such a large data. We tried to apply the kmeans algorithm provided in R, and it took more than 15 iteration for each metric.
2. There was a problem sshing into EMR because of which we could not run our code on the cluster, but we are hoping to fix it.

